---
title: "Week 3 Lab: Sentiment Analysis I"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

# Assignment

0. Using the “IPCC” Nexis Uni data set from the class presentation and the pseudo code we discussed, recreate Figure 1A from Froelich et al. (Date x # of 1) positive, 2) negative, 3) neutral headlines):

1.  Access the Nexis Uni database through the UCSB library: <https://www.library.ucsb.edu/research/db/211>

2.  Choose a key search term or terms to define a set of articles. 

3.  Use your search term along with appropriate filters to obtain and download a batch of at least 100 full text search results (.docx).  

4.  Read your Nexis article document into RStudio and parse with pdftools.

5.  This time use the full text of the articles for the analysis.  First clean any artifacts of the data collection process (hint: this type of thing should be removed: "Apr 04, 2022( Biofuels Digest: http://www.biofuelsdigest.com/ Delivered by Newstex"))

6.  Explore your data a bit and try to replicate some of the analyses above presented in class if you'd like (not necessary).  

7. Plot the amount of emotion words (the 8 from nrc) as a percentage of all the emotion words used each day (aggregate text from articles published on the same day).  How does the distribution of emotion words change over time?  Can you think of any reason this would be the case?

```{r eval=TRUE}
library(tidyr) #text analysis in R
library(lubridate) #working with date data
library(pdftools) #read in pdfs
library(tidyverse)
library(tidytext)
library(here)
library(LexisNexisTools) #Nexis Uni data wrangling
library(sentimentr)
library(readr)
```
## Set up Sentiment Words
```{r nrc_sentiment}

nrc_sent <- get_sentiments('nrc') #requires downloading a large dataset via prompt

nrc_fear <- get_sentiments("nrc") %>% 
  filter(sentiment == "fear")
```

## Key Search Term = decarbonization

- Search term = decarbonization
- Narrow by: news and English language

## Download Batch of data and Read in

```{r nexis_data, message=FALSE}
my_files <- list.files(pattern = ".docx", path = here("data/decarbonization/"),
                       full.names = TRUE, recursive = TRUE, ignore.case = TRUE)

dat <- lnt_read(my_files) #Object of class 'LNT output'

# lets make the tibbles in the LNToutput into individual df

meta_df <- dat@meta
articles_df <- dat@articles
paragraphs_df <- dat@paragraphs

dat2<- data_frame(element_id = seq(1:length(meta_df$Headline)), Date = meta_df$Date, Headline = meta_df$Headline)

paragraphs_dat <- data_frame(element_id = paragraphs_df$Art_ID, Text  = paragraphs_df$Paragraph)

dat3 <- inner_join(dat2,paragraphs_dat, by = "element_id")

```

## Use the full text of the articles for the analysis.  First clean any artifacts of the data collection process (hint: this type of thing should be removed: "Apr 04, 2022( Biofuels Digest: http://www.biofuelsdigest.com/ Delivered by Newstex"))

## Explore your data a bit and try to replicate some of the analyses above presented in class if you'd like (not necessary).  

```{r}
#most common words by sentiment
fear_words <- dat3  %>%
  unnest_tokens(output = word, input = Text, token = 'words') %>%
  inner_join(nrc_fear) %>%
  count(word, sort = TRUE)
head(fear_words)
```

```{r}
#unnest to word-level tokens, remove stop words, and join sentiment words
text_words <- dat3  %>%
  unnest_tokens(output = word, input = Text, token = 'words')

nrc_word_counts <- text_words %>%
  inner_join(get_sentiments("nrc")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

head(nrc_word_counts)
```

Let's break it out and plot the contributions by particular words different sentiment categories
```{r sent_counts}
decarb_sent_counts <- text_words %>%
        group_by(element_id) %>%
        # mutate(page_num = 1:n(),
        #        index = round(page_num / n(), 2)) %>%
        #unnest_tokens(word, line) %>%
        inner_join(get_sentiments("nrc")) %>%
        group_by(sentiment) %>%
        count(word, sentiment, sort = TRUE) %>%
        ungroup()

decarb_sent_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```

## Plot the amount of emotion words (the 8 from nrc) as a percentage of all the emotion words used each day (aggregate text from articles published on the same day).  How does the distribution of emotion words change over time?  Can you think of any reason this would be the case?

#### edit below
```{r}
#can we create a similar graph to Figure 3A from Froelich et al.? 
mytext <- get_sentences(dat3$Headline)
sent <- sentiment(mytext)

sent_df <- inner_join(dat3, sent, by = "element_id")

sentiment <- sentiment_by(sent_df$Headline)

sent_df %>%
  arrange(sentiment)

```

Pseudo code for Froelich et al plot:
mean sentiment by day, summarize
plot x = day, y = sentiment
```{r}

```


```{r custom_stop_words}
custom_stop_words <- bind_rows(tibble(word = c("your_word"),  
                                      lexicon = c("custom")), 
                               stop_words)
```



## Assignment (Due 4/19 by 5PM)

1.  Access the Nexis Uni database through the UCSB library: <https://www.library.ucsb.edu/research/db/211>

2.  Choose a key search term or terms to define a set of articles. 

3.  Use your search term along with appropriate filters to obtain and download a batch of at least 100 full text search results (.docx).  

4.  Read your Nexis article document into RStudio and parse with pdftools.

5.  This time use the full text of the articles for the analysis.  First clean any artifacts of the data collection process (hint: this type of thing should be removed: "Apr 04, 2022( Biofuels Digest: http://www.biofuelsdigest.com/ Delivered by Newstex"))

6.  Explore your data a bit and try to replicate some of the analyses above presented in class if you'd like (not necessary).  

7. Plot the amount of emotion words (the 8 from nrc) as a percentage of all the emotion words used each day (aggregate text from articles published on the same day).  How does the distribution of emotion words change over time?  Can you think of any reason this would be the case?

