---
title: 'Topic 7: Word Embeddings'
author: "Joe DeCesaro"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This week's Rmd file here: <https://github.com/MaRo406/EDS_231-text-sentiment/blob/main/topic_7.Rmd>

```{r packages, include = FALSE}
library(here)
library(tidytext)
library(tidyverse)
library(widyr)
library(irlba) #singluar value decomposition
library(broom) # creating search_synonym function
library(textdata)
library(ggplot2)
library(dplyr)

#https://semantle.com/
```


### Assignment

Download a set of pretrained vectors, GloVe, and explore them. 

Grab data here:
<!-- download.file('<https://nlp.stanford.edu/data/glove.6B.zip>',destfile = 'glove.6B.zip')  -->
<!-- unzip('glove.6B.zip')  -->
<!-- Use this file: 'glove.6B.300d.txt' -->

Use the last three chunks of this markdown to produce the assignment.

```{r}
wiki_data <- read_table(file = here('data/glove/glove.6B.300d.txt'),
                        col_names = FALSE)
wiki_data <- wiki_data %>% 
  column_to_rownames(var = "X1")
#rownames(wiki_data) <- wiki_data$X1

word_vectors <- as.matrix(x = wiki_data)
```

```{r}
search_synonyms <- function(word_vectors, selected_vector) {
dat <- word_vectors %*% selected_vector
    
similarities <- dat %>%
        tibble(token = rownames(dat), similarity = dat[,1])

similarities %>%
       arrange(-similarity) %>%
        select(c(2,3))
}
```

1.  Recreate the analyses in the last three chunks (find-synonyms, plot-synonyms, word-math) with the GloVe embeddings. How are they different from the embeddings created from the climbing accident data? Why do you think they are different?

```{r}
fall <- search_synonyms(word_vectors,word_vectors["fall",])
slip <- search_synonyms(word_vectors,word_vectors["slip",])
```

```{r}
slip %>%
    mutate(selected = "slip") %>%
    bind_rows(fall %>%
                  mutate(selected = "fall")) %>%
    group_by(selected) %>%
    top_n(15, similarity) %>%
    ungroup %>%
    mutate(token = reorder(token, similarity)) %>%
    ggplot(aes(token, similarity, fill = selected)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~selected, scales = "free") +
    coord_flip() +
    theme(strip.text=element_text(hjust=0, size=12)) +
    scale_y_continuous(expand = c(0,0)) +
    labs(x = NULL, title = "What word vectors are most similar to slip or fall?")
         
```

These graphs vary wildly from the climbing incident data with words close to fall being much more associated with financial words or closer to the word itself like "falling". Slip also has much similar words, like "slips", but also seems to have a greater variety of similar words. We did not remove variations of words in this data so that is why we are getting slips, falling, and more. The climbing data set was for the sport so it makes sense that there are different word associations when compared to this data.

```{r}
# take semantics of snow and danger and see what happens when they are added together
snow_danger <- word_vectors["snow",] + word_vectors["danger",] 
search_synonyms(word_vectors, snow_danger)

# remove snow and association of snow from danger and see what happens
no_snow_danger <- word_vectors["danger",] - word_vectors["snow",] 
search_synonyms(word_vectors, no_snow_danger)
```

Snow and danger together seems to have a lot more weather words than in the climbing data. When snow association is removed from danger it seems to focus on risk and some other, more random words.

2.  Run the classic word math equation, "king" - "man" = ?
```{r}
no_king_man <- word_vectors["king",] - word_vectors["man",] 
search_synonyms(word_vectors, no_king_man)
```

We get a lot of words that are likely the word "king" in other languages or names of kings.

3.  Think of three new word math equations. They can involve any words you'd like, whatever catches your interest.
```{r}
no_baseball_bat <- word_vectors["baseball",] - word_vectors["bat",] 
search_synonyms(word_vectors, no_baseball_bat)
```

```{r}
no_surfing_wave <- word_vectors["surfing",] - word_vectors["wave",] 
search_synonyms(word_vectors, no_surfing_wave)
```

```{r}
santa_barbara <- word_vectors["santa",] + word_vectors["barbara",] 
search_synonyms(word_vectors, santa_barbara)
```